Reliability Demonstration of A Novel Randomized Measurement and Verification Method for Switchable Control Retrofit Using Large Open Source Dataset

### 2025-02-01 code/manuscript.Rmd
changed interval dataset error and modified graph, manuscript

---
title: "Reliability Demonstration of A Novel Randomized Measurement and Verification Method for Switchable Control Retrofit Using Large-scale Public Dataset"
author:
  - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA], Paul Raftery ^1^, Stefano Schiavon ^1^, Carlos Durate ^1^
abstract: Conventional measurement and verification (M&V) methods relying on pre- and post-  retrofit comparison for estimating whole-building energy savings are often time-consuming and unreliable, especially when non-routine events such as reduced occupancy occur during the M&V process. Those events are unrelated to the intervention but significantgly affect building energy consumption and thus when the analyst applies conventiona M&V suggested by industry guidelines, the results can be largely confounded. In this study, we demonstrated that switchable interventions, such as most of the control retrofits, can benefit from random sampling, where the analyst randomly selects whether to implement the baseline or the intervention strategy each day. We tested this novel randomized M&V method on a large public dataset covering various climate zones and commercial building types, using a virtual chilled water supply temperature reset based on outdoor weather as the intervention. Our results show that comparing to the conventional method, the randomized method provides faster and more accurate savings estimation. Additionaly we found that the randomized M&V approach estimates much closer to the true savings when non-routine events are present demonstrating much improved reliability.

@@

# load packages using pacman
pacman::p_load(tidyverse, lubridate, here, stats, zoo, scales, ggpubr, patchwork, RColorBrewer)

@@

# colors
ls_colors <- c("Baseline" = "#c6dbef",
               "Measured baseline" = "#c6dbef",
               "Adjusted baseline" = "#c6dbef",
               "Projected baseline" = "#2171b5",
               "Intervention" = "#fdbb84",
               "Measured interv" = "#fdbb84",
               "Conventional M&V" = "grey70", 
               "Randomized M&V" = "#66c2a4", 
               "Buildings satisfying all criteria" = "#66c2a4", 
               "Buildings satisfying 80% TMY range" = "black", 
               "Buildings satisfying SPRT" = "black")

@@

# Tidy dataset
readfile_tidy <- str_glue("../readfiles/tidy/")
fig_path = "../figs/manuscript/"

df_energy_tidy <- read_rds(paste0(readfile_tidy, "df_energy.rds"))
df_meta_tidy <- read_rds(paste0(readfile_tidy, "df_meta.rds"))
df_weather_tidy <- read_rds(paste0(readfile_tidy, "df_weather.rds"))
df_sprt_all_tidy <- read_rds(paste0(readfile_tidy, "df_sprt_all.rds"))
df_seq_FS_tidy <- read_rds(paste0(readfile_tidy, "df_seq_FS.rds"))
df_NRE_occ <- read_rds(paste0(readfile_tidy, "df_NRE_occ.rds"))
df_MD_tidy <- read_rds(paste0(readfile_tidy, "df_MD.rds"))
df_FS_tidy <- read_rds(paste0(readfile_tidy, "df_FS.rds"))
df_eui_tidy <- read_rds(paste0(readfile_tidy, "df_eui.rds"))
df_cont_MD_tidy <- read_rds(paste0(readfile_tidy, "df_cont_MD.rds"))
df_cont_FS_tidy <- read_rds(paste0(readfile_tidy, "df_cont_FS.rds"))
df_model_acc_tidy <- read_rds(paste0(readfile_tidy, "df_model_acc.rds"))

all_sites_tidy <- df_energy_tidy %>%
  select(site) %>%
  distinct() %>%
  arrange(site)

all_types_tidy <- df_energy_tidy %>%
  select(type) %>%
  mutate(type = as.factor(type)) %>%
  distinct()

all_names_tidy <- df_energy_tidy %>%
  select(name) %>%
  distinct(name)

# Messy dataset
readfile_messy <- str_glue("../readfiles/messy/")

df_energy_messy <- read_rds(paste0(readfile_messy, "df_energy.rds"))
df_meta_messy <- read_rds(paste0(readfile_messy, "df_meta.rds"))
df_weather_messy <- read_rds(paste0(readfile_messy, "df_weather.rds"))
df_sprt_all_messy <- read_rds(paste0(readfile_messy, "df_sprt_all.rds"))
df_seq_FS_messy <- read_rds(paste0(readfile_messy, "df_seq_FS.rds"))
df_MD_messy <- read_rds(paste0(readfile_messy, "df_MD.rds"))
df_FS_messy <- read_rds(paste0(readfile_messy, "df_FS.rds"))
df_eui_messy <- read_rds(paste0(readfile_messy, "df_eui.rds"))
df_cont_MD_messy <- read_rds(paste0(readfile_messy, "df_cont_MD.rds"))
df_cont_FS_messy <- read_rds(paste0(readfile_messy, "df_cont_FS.rds"))
df_FS_nsprt <- read_rds(paste0(readfile_messy, "df_FS_nsprt.rds"))
df_MD_nsprt <- read_rds(paste0(readfile_messy, "df_MD_nsprt.rds"))
df_seq_FS_nsprt <- read_rds(paste0(readfile_messy, "df_seq_FS_nsprt.rds"))
df_model_acc_messy <- read_rds(paste0(readfile_messy, "df_model_acc.rds"))

all_sites_messy <- df_energy_messy %>%
  select(site) %>%
  distinct() %>%
  arrange(site)

all_types_messy <- df_energy_messy %>%
  select(type) %>%
  mutate(type = as.factor(type)) %>%
  distinct()

all_names_messy <- df_energy_messy %>%
  select(name) %>%
  distinct(name)

# read functions
function_path <- "../functions/"
source(paste0(function_path, "model_fit.R"))
source(paste0(function_path, "model_pred.R"))
source(paste0(function_path, "prepost_plot.R"))
```

@@

For the purpose of this study, we only queried whole-building hourly
electricity usage. All building names and precise locations were erased
but the climate zone and city name were provided, which are shown in
Figure \@ref(fig:tidysite) and Figure \@ref(fig:messysite). In addition,
the measurements were pre-processed by the authors with timestamp
already converted to local time. We further described the data filtering
process in Section \@ref(method).

@@

### Measurement and verification energy-efficient measures

@@

### Impact of non-routine events on building energy usage

@@

### Randomized experimental design in other scientific fields

To address the research gap identified in existing M&V studies,
particularly the challenges posed by non-routine events, we applied a
randomized experimental design method---an approach commonly used in
other scientific disciplines. In clinical trials, for instance,
randomization is employed to determine the effect of a medical treatment
(such as a new drug) by randomly assigning participants to either a
control group (e.g., receiving a placebo) or a treatment group
[@angus_adaptive_2019, @burger_importance_2021, @lim_randomization_2019,
@wiley_crossover_2016]. The primary purpose of randomization is to block
confounding variables such as age, gender, and socioeconomic status when
evaluating treatment effects. In the context of buildings, a "treatment"
refers to a control retrofit, while the "placebo" represents the
existing baseline control. Unlike clinical trials, where participants
are randomly assigned, our approach randomizes treatment assignment
longitudinally for each individual building---an approach known as
n-of-1 trials [@gabler2011]. Although the BDG2 dataset allows for
randomization at a population level (e.g., across climate zones),
building owners are generally more interested in understanding the
effects of control retrofits specific to their own buildings, rather
than at a generalized population level.

@@

## Building filtering

### 'Tidy' subset

In this study, we extracted the electricity measurements from the BDG2
dataset. On a first pass, we filtered out buildings with less noise
based on the following criteria:

@@

3.  No statistical significant difference (P-value \> 0.05) between the
    two-year electricity usage: target buildings should have no change
    in the electricity usage between the two years.

4.  Target buildings should have known site location: buildings with
    anonymous location are excluded due to unavailable typical
    meteorological weather.

5.  Warehouse and parking types are excluded: target buildings have less
    demand flexibility to implement a chilled water setpoint reset
    control.

Therefor the resulting subset contains all buildings with 'tidy'
measurements. Figure \@ref(fig:tidysite) shows in total, the subset
contains `r nrow(all_names_tidy)` buildings of `r nrow(all_types_tidy)`
types.

```{r tidysite, fig.cap = "Site summary of the tidy building subset (counts < 2 are omitted for visualization; left: aggragated counts of buildings for each type; right: breakdown building counts for each building type at each location)", fig.width=8, fig.height=6}

set3 <- colorRampPalette(brewer.pal('Set3',n=12))
type_colors <- setNames(set3(13), all_types_messy$type)

p1 <- df_energy_tidy %>%
  group_by(type) %>%
  distinct(name) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  mutate(type = fct_reorder(type, n, .desc = F)) %>%
  ggplot(aes(x = 1, y = n, fill = as.factor(type))) +
  geom_col() +
  geom_text(aes(label = ifelse(n > 2, as.character(n), "")), color = "black", position = position_stack(vjust = 0.5)) +
  scale_y_continuous(expand = c(0, 0),
                     breaks = breaks_pretty(n = 4)) +
  scale_x_discrete(expand = c(0, 0.1)) +
  scale_fill_manual(values = type_colors) +
  labs(x = NULL,
       y = NULL,
       subtitle = "Combining all locations",
       fill = NULL) +
  theme(axis.text.x = element_blank(),
        panel.grid.major.y = element_line(color = "grey80"),
        legend.direction = "horizontal",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

  p2 <- df_energy_tidy %>%
    left_join(df_loc, by = "site") %>% 
    group_by(location, type) %>%
    distinct(name) %>%
    summarise(n = n()) %>%
    mutate(proportion = n / sum(n),
           total = sum(n),
           ymax = cumsum(proportion),
           ymin = c(0, head(ymax, n = -1))) %>%
    mutate(label_pos = (ymax + ymin) / 2) %>%
    ungroup() %>%
    group_by(type) %>%
    mutate(order = sum(n)) %>%
    ungroup() %>%
    mutate(type = fct_reorder(type, order, .desc = F)) %>%
    ggplot(aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 3, fill = type)) +
    geom_rect() +
    coord_polar(theta = "y") +
    xlim(c(2, 4)) +
    facet_wrap(~ location, nrow = 3) +
    labs(x = NULL,
         y = NULL,
         subtitle = "For each location",
         fill = NULL) +
    scale_fill_manual(values = type_colors) +
    geom_text(aes(x = 3.5, y = label_pos, label = ifelse(n > 2, as.character(n), ""))) +
    geom_text(aes(x = 2, y = 0, label = paste0("Total\n", total)), color = "black") +
    theme(legend.direction = "horizontal",
          axis.text = element_blank(),
          plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

  ggarrange(p1, p2,
            ncol = 2, nrow = 1,
            widths = c(0.2, 1),
            common.legend = TRUE,
            legend = "bottom") +
    plot_annotation(title = "Summary of building numbers and types for 'tidy' subset")
```

### 'Messy' subset

In reality, the 'tidy' subset is less representative of the measurements
typically collected from existing building stock, as whole-building
electricity measurements collected over two years often show more
variability the the filtering critieron. These variations can stem from
issues like sensor calibration errors or inherent changes in the
building, which, as previously discussed, can bias M&V results.
Therefore, to more realistically assess the robustness of the two M&V
method, we included an additional 'messy' subset which first exclude the
'tidy' subset and then re-apply the filtering rule with one amendment:

3.  $\frac{abs(E_{2017} - E_{2016})}{E_{2016}}< 25\%$: any increase or
    decrease of building electricity usage in the second year should be
    less than 25% of that in th first year.

Figure \@ref(fig:messysite) shows the summary of the 'messy' dataset,
which contains `r nrow(all_names_messy)` buildings in
`r nrow(all_types_messy)` types.

```{r messysite, fig.cap = "Site summary of the messy building subset (counts < 2 are omitted for visualization; left: aggragated counts of buildings for each type; right: breakdown building counts for each building type at each location)", fig.width=8, fig.height=8}

p1 <- df_energy_messy %>%
    group_by(type) %>%
    distinct(name) %>%
    summarise(n = n()) %>%
    ungroup() %>%
    mutate(type = fct_reorder(type, n, .desc = F)) %>%
    ggplot(aes(x = 1, y = n, fill = as.factor(type))) +
    geom_col() +
    geom_text(aes(label = ifelse(n > 10, as.character(n), "")), color = "black", position = position_stack(vjust = 0.5)) +
    scale_y_continuous(expand = c(0, 0),
                       breaks = breaks_pretty(n = 4)) +
    scale_x_discrete(expand = c(0, 0.1)) +
    scale_fill_manual(values = type_colors) +
    labs(x = NULL,
         y = NULL,
         subtitle = "Combining all locations",
         fill = NULL) +
    theme(axis.text.x = element_blank(),
          panel.grid.major.y = element_line(color = "grey80"),
          legend.direction = "horizontal",
          plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

  p2 <- df_energy_messy %>%
    left_join(df_loc, by = "site") %>% 
    group_by(location, type) %>%
    distinct(name) %>%
    summarise(n = n()) %>%
    mutate(proportion = n / sum(n),
           total = sum(n),
           ymax = cumsum(proportion),
           ymin = c(0, head(ymax, n = -1))) %>%
    mutate(label_pos = (ymax + ymin) / 2) %>%
    ungroup() %>%
    group_by(type) %>%
    mutate(order = sum(n)) %>%
    ungroup() %>%
    mutate(type = fct_reorder(type, order, .desc = F)) %>%
    ggplot(aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 3, fill = type)) +
    geom_rect() +
    coord_polar(theta = "y") +
    xlim(c(2, 4)) +
    facet_wrap(~ location, nrow = 3) +
    labs(x = NULL,
         y = NULL,
         subtitle = "For each location",
         fill = NULL) +
    scale_fill_manual(values = type_colors) +
    geom_text(aes(x = 3.5, y = label_pos, label = ifelse(n > 1, as.character(n), ""))) +
    geom_text(aes(x = 2, y = 0, label = paste0("Total\n", total)), color = "black") +
    theme(legend.direction = "horizontal",
          axis.text = element_blank(),
          plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

  ggarrange(p1, p2,
            ncol = 2, nrow = 1,
            widths = c(0.2, 1),
            common.legend = TRUE,
            legend = "bottom") +
    plot_annotation(title = "Summary of building numbers and types for 'messy' subset")
```

## Apply control intervention

@@

## Apply example non-routine event

To illustrate the impact of non-routine events that are not properly
adjusted for, we developed three scenarios to quantify the energy
consumption changes associated with a hypothetical increase in
occupancy. Specifically, we assume that a rise in occupancy in 2016
(referred to as the "baseline year" when using the conventional method)
led to a 20% increase in whole-building measured electricity usage. The
three scenarios describe when this increase occurred: January to April
(S1), May to August (S2), and September to December (S3). This
hypothetical change was applied only to buildings in the "tidy" subset,
as these buildings were initially more likely to be free from
non-routine events.

## Run M&V methods

As described in Section \@ref(introduction), conventional method for M&V
is time-consuming and unable to separate non-routine event impact from
measured savings. If using conventional method to estimate intervention
energy savings, the result (after 24 months) is calculated by the
difference between measured intervention and projected baseline in the
post-retrofit period. In this study, we leveraged a piece-wise linear
regression considering time-of-week and outdoor temperature (TOWT) as
independent variables for projection for the counter-factual baseline in
the post-retrofit period and normalization on typical meteorological
year [@mathieu_quantifying_2011].

Compared to the conventional method, the randomized M&V approach offers
a more rapid and reliable estimation of energy savings. To apply this
method, analysts first define the target savings and stopping criteria.
Then they design a randomized switchback schedule and perform sequential
statistical tests, such as the sequential probability ratio test (SPRT),
to monitor savings as data is collected. Once the target savings are
detected, the analysts fit a prediction model (e.g. TOWT) to adjust for
differences in outdoor temperature, following the same adjustment
process as the conventional method. We provide example switchback
experimental design and stopping criteria thresholds below:

@@

Stopping criteria are:

@@

To remain consistency, savings normalization on typical meteorological
year is also modeled through TOWT.

@@

In this section, we compare the performance of two M&V methods. The key
aspects of the assessment include: 1) time required to reach a saving
estimation: in most cases, a shorter M&V timeline reduces associated
cost and interruption for the building owner; and 2) saving estimation
accuracy: this is particularly important for any software-as-a-service
company to set a reasonable price with their customers.

### Savings estimation time

ASHRAE Guideline 14 offering minimum requirements for whole-building
measurement path states that the baseline period is either a full range
of all independent variables (typically outdoor weather conditions)
under normal facility operation or a 12-month worth of continuous
measurements. The same requirements also applies to intervention
installed in the post-retrofit period. Thus, the conventional M&V method
is likely to run over 24 months or even longer due to missing data. For
example if an analyst is asked to determine the energy savings of a
chilled water plant retrofit but missed most of the cooling season due
to delay in retrofit deployment, he/she needs to wait until the next
cooling season to measure the savings. If the randomized M&V method is
applicable, delays in retrofit deployment pose less risk to the building
owner since baseline are continuously monitored. Sampling at equal
probabilities (e.g., 50%/50% between baseline and intervention) helps to
balance the distribution of independent variables like outdoor weather
conditions across all sampled control strategies. As long as one
strategy is not sampled over a large number of consecutive days (e.g., 7
days or more), it is unlikely that the measurements of independent
variables will differ significantly between the control strategies.

Figure \@ref(fig:timeline) shows the average estimated timeline for each
building across all sites and climate conditions if using randomized M&V
for the applied chilled water supply temperature reset intervention. The
timeline figures for each individual building are attached in the
supplementary material. The figure shows all buildings can detect a
saving statistically from sampled measurements within 12 months, which
is equivalent to the minimum test criterion. Similar to the findings
from our pervious study, covering a sufficient range of outdoor weather
condition is the most stringent requirement. If target buildings are
located in a climate zone with more preferable outdoor weather
conditions such as California, covering 80% of the TMY range should only
take 3 \~ 4 months. In this case, the analyst can conclude the M&V
simultaneously if the blocking period ends at the same time. However,
given most buildings requires 6 months to achieve 80% of the TMY range,
all buildings can determine the savings, including associated
uncertainties, within 9 months---significantly shorter than the baseline
measurement period required by the conventional method.

```{r timeline, fig.cap="Randomized M&V timeline of satisfying key stopping criteria for all buildings", fig.width=8, fig.height=4}

df_time <- df_sprt_all_messy %>% 
  filter(seq != "final") %>% 
  bind_rows(df_sprt_all_tidy %>% filter(seq != "final")) %>% 
  select(name, seq, n_weeks)

count <- list()

for (i in seq(1, 36, by = 1)){

>>>>

---
title: "Reliability Demonstration of A Novel Randomized Measurement and Verification Method for Switchable Control Retrofit Using Large Open Source Dataset"
author:
  - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA], Paul Raftery ^1^, Stefano Schiavon ^1^, Carlos Duarte ^1^, Gail Brager ^1^
abstract: Conventional measurement and verification (M&V) methods relying on pre- and post-  retrofit comparison for estimating whole-building energy savings are often time-consuming and unreliable, especially when non-routine events such as reduced occupancy occur during the M&V process. Those events are unrelated to the intervention but significantgly affect building energy consumption and thus when the analyst applies conventiona M&V suggested by industry guidelines, the results can be largely confounded. In this study, we demonstrated that switchable interventions, such as most of the control retrofits, can benefit from random sampling, where the analyst randomly selects whether to implement the baseline or the intervention strategy each day. We tested this novel randomized M&V method on a large public dataset covering various climate zones and commercial building types, using a virtual chilled water supply temperature reset based on outdoor weather as the intervention. Our results show that comparing to the conventional method, the randomized method provides faster and more accurate savings estimation. Additionaly we found that the randomized M&V approach estimates much closer to the true savings when non-routine events are present demonstrating much improved reliability.

@@

# load packages using pacman
pacman::p_load(tidyverse, lubridate, here, stats, zoo, scales, lvplot, ggpubr, gridExtra, patchwork, RColorBrewer)

@@

# colors
ls_colors <- c("Baseline" = "#c6dbef",
               "Measured baseline" = "#c6dbef",
               "Adjusted baseline" = "#c6dbef",
               "Projected baseline" = "#2171b5",
               "Intervention" = "#fdbb84",
               "Measured interv" = "#fdbb84",
               "Conventional" = "grey70", 
               "Randomized" = "#99d8c9", 
               "Randomized\n(24 months)" = "#66c2a4",
               "Randomized\n(50/50)" = "#66c2a4",
               "Daily\nsampling" = "#66c2a4",
               "2-day\nsampling" = "#66c2a4",
               "3-day\nsampling" = "#66c2a4",
               "7-day\nsampling" = "#66c2a4",
               "Dropped" = "#41ae76",
               "Kept" = "#006d2c",
               "Randomized\n(20/80)" = "#ccece6",
               "Buildings finishing randomized M&V" =  "#99d8c9",
               "Buildings satisfying 80% TMY range" = "black", 
               "Buildings satisfying SPRT" = "black")

@@

# stable dataset
readfile_stable <- str_glue("../readfiles/stable/")
fig_path = "../figs/manuscript/"

df_energy_stable <- read_rds(paste0(readfile_stable, "df_energy.rds"))
df_weather_stable <- read_rds(paste0(readfile_stable, "df_weather.rds"))
df_sprt_all_stable <- read_rds(paste0(readfile_stable, "df_sprt_all.rds"))
df_seq_fs_stable <- read_rds(paste0(readfile_stable, "df_seq_fs.rds"))
df_fs_stable <- read_rds(paste0(readfile_stable, "df_fs.rds"))
df_cont_stable <- read_rds(paste0(readfile_stable, "df_cont.rds"))
df_fs_null_stable <- read_rds(paste0(readfile_stable, "df_fs_null.rds"))
df_seq_fs_null_stable <- read_rds(paste0(readfile_stable, "df_seq_fs_null.rds"))
df_model_acc_stable <- read_rds(paste0(readfile_stable, "df_model_acc.rds"))
df_fs_tmy_stable <- read_rds(paste0(readfile_stable, "df_fs_tmy.rds"))
df_fs_tmy_null_stable <- read_rds(paste0(readfile_stable, "df_fs_tmy_null.rds"))
df_interval_drop_stable <- read_rds(paste0(readfile_stable, "df_interval_drop.rds"))
df_interval_keep_stable <- read_rds(paste0(readfile_stable, "df_interval_keep.rds"))
df_seq_interval_fs_drop_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_fs_drop.rds"))
df_seq_interval_fs_keep_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_fs_keep.rds"))
df_seq_interval_nm_drop_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_nm_drop.rds"))
df_seq_interval_nm_keep_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_nm_keep.rds"))
df_seq_interval_tl_drop_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_tl_drop.rds"))
df_seq_interval_tl_keep_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_tl_keep.rds"))

all_sites_stable <- df_energy_stable %>%
  select(site) %>%
  distinct() %>%
  arrange(site)

all_types_stable <- df_energy_stable %>%
  select(type) %>%
  mutate(type = as.factor(type)) %>%
  distinct()

all_names_stable <- df_energy_stable %>%
  select(name) %>%
  distinct(name)

# variable dataset
readfile_variable <- str_glue("../readfiles/variable/")

df_energy_variable <- read_rds(paste0(readfile_variable, "df_energy.rds"))
df_weather_variable <- read_rds(paste0(readfile_variable, "df_weather.rds"))
df_sprt_all_variable <- read_rds(paste0(readfile_variable, "df_sprt_all.rds"))
df_seq_fs_variable <- read_rds(paste0(readfile_variable, "df_seq_fs.rds"))
df_fs_variable <- read_rds(paste0(readfile_variable, "df_fs.rds"))
df_cont_variable <- read_rds(paste0(readfile_variable, "df_cont.rds"))
df_fs_null_variable <- read_rds(paste0(readfile_variable, "df_fs_null.rds"))
df_seq_fs_null_variable <- read_rds(paste0(readfile_variable, "df_seq_fs_null.rds"))
df_model_acc_variable <- read_rds(paste0(readfile_variable, "df_model_acc.rds"))
df_fs_tmy_variable <- read_rds(paste0(readfile_variable, "df_fs_tmy.rds"))
df_fs_tmy_null_variable <- read_rds(paste0(readfile_variable, "df_fs_tmy_null.rds"))
df_interval_drop_variable <- read_rds(paste0(readfile_variable, "df_interval_drop.rds"))
df_interval_keep_variable <- read_rds(paste0(readfile_variable, "df_interval_keep.rds"))
df_seq_interval_fs_drop_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_fs_drop.rds"))
df_seq_interval_fs_keep_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_fs_keep.rds"))
df_seq_interval_nm_drop_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_nm_drop.rds"))
df_seq_interval_nm_keep_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_nm_keep.rds"))
df_seq_interval_tl_drop_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_tl_drop.rds"))
df_seq_interval_tl_keep_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_tl_keep.rds"))

all_sites_variable <- df_energy_variable %>%
  select(site) %>%
  distinct() %>%
  arrange(site)

all_types_variable <- df_energy_variable %>%
  select(type) %>%
  mutate(type = as.factor(type)) %>%
  distinct()

all_names_variable <- df_energy_variable %>%
  select(name) %>%
  distinct(name)

# read functions
function_path <- "../functions/"
source(paste0(function_path, "model_fit.R"))
source(paste0(function_path, "model_pred.R"))
source(paste0(function_path, "prepost_plot.R"))

S_building <- nrow(df_fs_tmy_stable)
V_building <- nrow(df_fs_tmy_variable)
A_building <- S_building + V_building
```

@@

@@

### Whole building approach

@@

### Non-routine events impact

@@

@@

## Data filtering

In this study, we extracted the electricity measurements from the BDG2
dataset and filter out all qualified buildings based on the following
criteria:

@@

3.  $\frac{abs(E_{2017} - E_{2016})}{E_{2016}}< 25\%$: any increase or
    decrease of building electricity usage in the second year should be
    less than 25% of that in the first year.

4.  Warehouse and parking types are excluded: target buildings have less
    demand flexibility to implement a chilled water setpoint reset
    control.

```{r stablesite, fig.cap = "Site summary of the stable building subset (counts < 2 are omitted for visualization; left: aggragated counts of buildings for each type; right: breakdown building counts for each building type at each location)", fig.width=8, fig.height=6}

set3 <- colorRampPalette(brewer.pal('Set3',n=12))
type_colors <- setNames(set3(13), all_types_variable$type)

stable_set <- df_energy_stable %>%
  group_by(type) %>%
  distinct(name) %>%
  summarise(stable = n()) %>%
  ungroup()

variable_set <- df_energy_variable %>%
  group_by(type) %>%
  distinct(name) %>%
  summarise(variable = n()) %>%
  ungroup()

p1 <- variable_set %>% 
  left_join(stable_set, by = "type") %>% 
  mutate(stable = replace_na(stable, 0)) %>% 
  group_by(type) %>% 
  summarise(n = stable + variable) %>% 
  ungroup() %>% 
  mutate(type = fct_reorder(type, n, .desc = F)) %>%
  ggplot(aes(x = 1, y = n, fill = as.factor(type))) +
  geom_col() +
  geom_text(aes(label = ifelse(n > 10, as.character(n), "")), color = "black", position = position_stack(vjust = 0.5)) +
  scale_y_continuous(expand = c(0, 0),
                     breaks = breaks_pretty(n = 4)) +
  scale_x_discrete(expand = c(0, 0.1)) +
  scale_fill_manual(values = type_colors) +
  labs(x = NULL,
       y = NULL,
       subtitle = "Combining all locations",
       fill = NULL) +
  theme(axis.text.x = element_blank(),
        panel.grid.major.y = element_line(color = "grey80"),
        legend.direction = "horizontal",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

stable_set <- df_energy_stable %>% 
  group_by(type, site) %>%
  distinct(name) %>%
  summarise(stable = n()) %>%
  ungroup()

variable_set <- df_energy_variable %>% 
  group_by(type, site) %>% 
  distinct(name) %>%
  summarise(variable = n()) %>%
  ungroup()

p2 <- variable_set %>% 
  left_join(stable_set, by = c("type", "site")) %>% 
  mutate(stable = replace_na(stable, 0)) %>% 
  group_by(site, type) %>% 
  summarise(n = stable + variable) %>% 
  ungroup() %>% 
  left_join(df_loc, by = "site") %>% 
  group_by(location) %>%
  mutate(proportion = n / sum(n),
         total = sum(n),
         ymax = cumsum(proportion),
         ymin = c(0, head(ymax, n = -1))) %>%
  mutate(label_pos = (ymax + ymin) / 2) %>%
  ungroup() %>%
  group_by(type) %>%
  mutate(order = sum(n)) %>%
  ungroup() %>%
  mutate(type = fct_reorder(type, order, .desc = F)) %>%
  ggplot(aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 3, fill = type)) +
  geom_rect() +
  coord_polar(theta = "y") +
  xlim(c(2, 4)) +
  facet_wrap(~ location, nrow = 3) +
  labs(x = NULL,
       y = NULL,
       subtitle = "For each location",
       fill = NULL) +
  scale_fill_manual(values = type_colors) +
  geom_text(aes(x = 3.5, y = label_pos, label = ifelse(n > 10, as.character(n), "")), check_overlap = T) +
  geom_text(aes(x = 2, y = 0, label = paste0("Total\n", total)), color = "black", check_overlap = T) +
  theme(legend.direction = "horizontal",
        axis.text = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

ggarrange(p1, p2,
          ncol = 2, nrow = 1,
          widths = c(0.2, 1),
          common.legend = TRUE,
          legend = "bottom") +
  plot_annotation(title = "Summary of building numbers and types")
```

For all qualified buildings, we further added one more stringent
criterion to label all buildings with very stable electricity usage
between the two years:

1.  No statistical significant difference (P-value \> 0.05) between the
    two-year electricity usage.

As a results, there are only 66 buildings were labeled in this 'stable'
subset (out of total of 600 buildings) implying that the assumption made
by conventional M&V is overly simplified. As mentioned in Section
\@ref(non-routine-events-impact), such variability is largely associated
with non-routine events and according to the statistics, is more
commonly observed in real applications.

## Control intervention implementation

@@

We described more in detail the workflow of both conventional and
randomized M&V methos in the previous study [@RAFTERY2024111134].

In a similar style, we summarized the experimental design as:

@@

and stopping criteria:

@@

@@

```{r abstract, fig.cap="Overall comparison results between the conventional M&V method and the proposed randomized M&V method (both at the stopping criteria and over a two-year period).", fig.height=12, fig.width=10}

#### GA ####
rand_eob_S <- df_seq_fs_stable %>% 
  filter(seq == "eob") %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - fs), 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_S <- df_fs_stable %>% 
  filter(method != "rand") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - conv), 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_S <- df_fs_stable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - rand), 
         method = "rand_final") %>% 
  select(name, method, diff)

df_MW_S <- rbind(rand_eob_S, conv_S, rand_final_S)

# plot for the variable subset
rand_eob_V <- df_seq_fs_variable %>% 
  filter(seq == "eob") %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site"))%>% 
  mutate(diff = abs(savings - fs), 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_V <- df_fs_variable %>% 
  filter(method != "rand") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - conv), 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_V <- df_fs_variable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - rand), 
         method = "rand_final") %>% 
  select(name, method, diff)

df_MW_V <- rbind(rand_eob_V, conv_V, rand_final_V)

# plot for combined dataset
rand_eob_A <- bind_rows(rand_eob_V, rand_eob_S) %>% 
  select(name, rand_eob_diff = diff)

conv_A <- bind_rows(conv_V, conv_S) %>% 
  select(name, conv_diff = diff)

rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
  select(name, rand_final_diff = diff)

df_MW_A <- rand_eob_A %>% 
  left_join(conv_A, by = "name") %>% 
  left_join(rand_final_A, by = "name") %>% 
  pivot_longer(c(rand_eob_diff, conv_diff, rand_final_diff), names_to = "method", values_to = "diff")

p_top <- df_MW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
  geom_text(data = . %>% group_by(method) %>% summarise(mean = mean(diff)) %>% ungroup(), 
            aes(x = method, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       color = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with measured weather")) +
  coord_cartesian(ylim = c(0, 23)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        axis.text.x = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

# TMY
# plot for the stable subset
rand_eob_S <- df_sprt_all_stable %>% 
  filter(seq == "eob") %>% 
  select(-c(seq, n_weeks)) %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - annual), 
         method = "rand_eob") %>% 
  select(name, method, diff)
